---
title: "Project title"
author: "Stephanie Christea, Nicole Sin, Johnson Vo, Lily Zuo, TUT107, group number: 1"
subtitle: The subtitle of my project
output:
  ioslides_presentation: default
  widescreen: yes
---


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# echo=FALSE will stop the code chunk from appearing in the knit document
# warning=FALSE and message=FALSE will stop R messages from appearing in the knit document
library(tidyverse)
```

---

Three dashes starts a new page when it is not started by a new header.

---

### This is a heading that doesn't start a new page

You can make text **bold** or in *italics*.

To make bullet points, start the points after a blank line:
 
- point one
- another point

---

## Read more

For more on creating an `ioslides` presentation see [https://rmarkdown.rstudio.com/ioslides_presentation_format.html](https://rmarkdown.rstudio.com/ioslides_presentation_format.html)

# Headings you should include in your project

## Introduction

Include here a few sentences to introduce the problem and provide context. You might want to briefly summarize the data in words (what is the data and what is it used for). You can present the questions you are investigating here. 

```{r}
neighbourhood <- auto_thefts %>% select(Index_, occurrencemonth, occurrencedayofweek, occurrencehour)%>% mutate(season = ifelse((occurrencemonth == 'March'| occurrencemonth == 'April'| occurrencemonth == 'May'| occurrencemonth == 'June'| occurrencemonth == 'July'| occurrencemonth == 'August'), yes = "Spring_Summer", no = "Fall_Winter"))%>%
  mutate(week_end_or_day = ifelse(occurrencedayofweek == "Sunday"| occurrencedayofweek == "Saturday", yes = "Weekend", no = "Weekday")) %>%
  mutate(time_of_day= ifelse(occurrencehour >= 18|occurrencehour <= 6, yes = "Night", no= "Day"))

glimpse(neighbourhood)
```

```{r}
#Histograms of occurrence month
neighbourhood %>% ggplot(aes(x = occurrencemonth)) + geom_bar() + ggtitle("Occurence Month")
#Histograms of occurrence day of week
neighbourhood %>% ggplot(aes(x = occurrencedayofweek)) + geom_bar() + ggtitle("Occurence Day of Week")
#Histograms of occurrence hour
neighbourhood %>% ggplot(aes(x = occurrencehour)) + geom_bar() + ggtitle("Occurence Hour")
```

```{r}
#Histograms of weekday and weekend
neighbourhood %>% 
  filter(!is.na(week_end_or_day)) %>%
  ggplot(aes(x = week_end_or_day)) + geom_bar(color = "black", fill = "pink") + ggtitle("Weekend vs Weekday")

```

```{r}
#simulation
n <- 18178 
repetitions <- 1000
sim <- rep(NA, repetitions)

set.seed(1071)
for (i in 1:repetitions) {
  sim_weekday_or_end <- sample(c("Weekday", "Weekend"),
       size = n,
       prob = c(5/7, 2/7),
       replace = TRUE) 
  sim_p <- sum(sim_weekday_or_end == "Weekday") / n
  sim[i] <- sim_p
}
#build a plot with simulated data (simulated distribution of weekday)
sim <- data_frame(p_weekdays = sim)
sim %>% ggplot(aes(x = p_weekdays)) +
  geom_histogram(binwidth = 0.001, bins = 30, colour = "black", fill = "pink") + ggtitle("Simulated Distribution of Weekdays") + labs(x = "Proportion weekdays in 18178 auto thefts" )
```

```{r}
#proportion of weekday/total (test statistic)
neighbourhood %>% filter(!is.na(week_end_or_day))%>%
  summarise(n(),
            n_weekday = sum(week_end_or_day == "Weekday"),
            prop = n_weekday / n()) 
```
```{r}
#test statistic
      13167/(13167+5008)
#     = 0.7244567

      5/7 - (0.7243371 - 5/7)
#     = 0.7042343
#p-value
      sim %>% filter(p_weekdays >= 0.7244567 | p_weekdays <= 0.7042343) %>%
        summarise(p_value = n() / repetitions)
#     = 0.004
```

```{r}
#histogram of datime and nighttime
neighbourhood %>% 
  filter(!is.na(time_of_day)) %>%
  ggplot(aes(x = time_of_day)) + geom_bar(color = "black", fill = "lightblue") + ggtitle("Daytime vs Nighttime")
```

```{r}
#simulation
n <- 18178
repetitions <- 1000
sim <- rep(NA, repetitions)

set.seed(1071)
for (i in 1:repetitions) {
  sim_time_of_day <- sample(c("Day", "Night"),
       size = n,
       prob = c(0.5, 0.5),
       replace = TRUE) 
  sim_p <- sum(sim_time_of_day == "Day") / n
  sim[i] <- sim_p
}
#build a plot with simulated data (simulated distribution of daytime)
sim <- data_frame(p_day = sim)
sim %>% ggplot(aes(x = p_day)) +
  geom_histogram(binwidth = 0.001, colour = "black", fill = "lightblue") + labs(x = "Proportion day time in 18178 auto thefts" )
```

```{r}
#proportion of day time (test statistic)
neighbourhood %>% filter(!is.na(time_of_day))%>%
  summarise(n(),
            n_daytime = sum(time_of_day == "Day"),
            prop = n_daytime / n())
            
```

```{r}
#test statistic
      6509/(18178)
#     = 0.3580702

      0.5 - (0.3580702 - 0.5)
#     = 0.6419298
#p-value
      sim %>% filter(p_day >= 0.6419298 | p_day <= 0.3580702) %>%
        summarise(p_value = n() / repetitions)
#     = 0

```


## Objectives (optional)

You can list the questions of interest in complete English sentences here to highlight them. 

## Data Summary (optional)

Here you can explain how you cleaned the data and created variables suitable for answering your questions. 
You can also include graphical displays that either motivated or address the questions.

## Statistical Methods

Describe here what you have done to the data without presenting any results (output). If you want to indicate variables by symbols or variable names, define them here. 

## Results

Present the main results here, in order of importance, related to the questions asked. You might use tables or graphs, or other ways to summarize your results.

## Conclusion

Give your main conclusions here. Follow the order of questions you presented. 

Here you can also mention any additional considerations, concerns, or issues you might have. For example, if the results were unexpected, you can discuss this and perhaps offer possible explanations.

## Acknowledgements (optional)

If you received any help from someone other than your team members you can acknowledge them. For example:   
*The authors would like to thank "TA name" for helpful suggestions and comments that improved the presentation of this poster.*

