---
title: "The Inclination for Auto Thefts in Relation to Time"
author: 'Stephanie Cristea, Nicole Sin, Johnson Vo, Lily Zuo, TUT107, Group Number:
  1'
subtitle: Based on Auto Thefts data from Toronto provided by the the Toronto Police
  Service and from the years 2014-2018
output:
  ioslides_presentation: default
  slidy_presentation: default
  transition: faster
  widescreen: yes
---


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# echo=FALSE will stop the code chunk from appearing in the knit document
# warning=FALSE and message=FALSE will stop R messages from appearing in the knit document
library(tidyverse)
library(rpart)
#install.packages("partykit")
library(partykit)
library(knitr)
auto_thefts <- read.csv("auto_thefts.csv")
neighbourhood <- auto_thefts %>% select(Index_, occurrencemonth, occurrencedayofweek, occurrencehour) %>% filter(occurrencemonth != '' & occurrencedayofweek != '') %>% mutate(season = ifelse((occurrencemonth == 'March'| occurrencemonth == 'April'| occurrencemonth == 'May'| occurrencemonth == 'June'| occurrencemonth == 'July'| occurrencemonth == 'August'), yes = "Spring_Summer", no = "Fall_Winter"))%>%
  mutate(week_end_or_day = ifelse(occurrencedayofweek == "Sunday    "| occurrencedayofweek == "Saturday  ", yes = "Weekend", no = "Weekday")) %>%
  mutate(time_of_day= ifelse(occurrencehour >= 18|occurrencehour <= 6, yes = "Night", no= "Day"))
```
## Goals

- Analyze whether time acts as a factor in the number of auto thefts - is there an inclination to when these crimes occur? 
  - Take an intricate look at the effect of the season and the hour on number of auto thefts
- Determine peak time when these crimes occur; when the public should take better precautions protecting their cars 
- Determine when police officers should be more prepared for auto-thefts based on when they take place more often


## Questions of Analysis

1. Does the proportion of auto thefts in the spring/summer differ from the proportion during fall/winter?
2. Does the number of autothefts show a significant correlation depending on whether it is during the day or the night?
3. Does the number of autothefts show a significant correlation depending on whether it is during the Spring/Summer or Winter/Fall?

# Data Visualization

---

```{r, echo=FALSE, warning=FALSE, fig.align = "center"}
neighbourhood %>% filter(!is.na(occurrencemonth)) %>% ggplot(aes(x = occurrencemonth, fill = season)) + geom_bar() + ggtitle("Number of Auto Thefts Depending on the Months \nof the Year") +
  labs (x = "Month", y = "Number of Auto Thefts")+coord_flip()+
  scale_fill_discrete(name="Seasons", breaks = c("Fall_Winter", "Spring_Summer"), labels=c("Fall and Winter", "Spring and Summer"))

```

---


```{r  echo=FALSE, warning=FALSE}
neighbourhood %>% ggplot(aes(x = occurrencehour, fill = time_of_day)) + geom_bar() + ggtitle("Number of Auto Thefts Depending on the Time \nof the Day")+
  labs (x = "Time of the day (by hour)", y = "Number of Auto Thefts")+
  scale_fill_discrete(name="Time of Day")
```


## Data Summary
The following variables were created to address our questions of analysis:

- A seasonal variable (categorizing an autotheft incidence as "Spring/ Summer" or "Fall/Winter")
- A day of the week variable (categorizing an autotheft incidence as "Weekday" or "Weekend") 
- A time of day variable (categorizing an autotheft incidence as "Day" or "Night")


## Question 1: 
Does the proportion of auto thefts in the spring/summer differ from the proportion during fall/winter?

```{r echo=FALSE, warning=FALSE}
neighbourhood %>%
  filter(!is.na(season))%>%
  ggplot(aes(x = season))+geom_bar(color = "black", fill = "lightblue")+
  ggtitle("Number of Autothefts based on Spring/Summer and Fall/Winter")+
  labs (x = "Spring/Summer vs Fall/Winter", y = "Number of Autothefts")
```



## Question 1 Results

**Method Used:**
- Simulation with a single variable was run 1000 times

**Results:**
- No evidence against our null hypothesis


```{r echo=FALSE, warning=FALSE}
# null hypothesis: the proportion of the incidence happened in spring and summer is 0.5.
# alternative hypothesis: the proportion of the incidence happened in spring and summer is not 0.5.
```


```{r echo=FALSE, warning=FALSE, eval = FALSE}
# test statistic
neighbourhood %>%
  filter(!is.na(season)) %>%
  summarise(n(),
            n_springsummer = sum (season == "Spring_Summer"),
            prop_springsummer = n_springsummer/n())
```

```{r echo=FALSE, warning=FALSE, eval = FALSE}
# simulation of incidence occurance in seasons. 
n_observations <- 18178
repetitions <- 1000
simulated_stats <- rep(NA, repetitions)
set.seed(1071)
for (i in 1:repetitions)
{
new_sim <- sample(c("Spring_Summer", "Fall_Winter"),
size = n_observations, prob=c(0.5,0.5), replace = TRUE)
sim_p <- sum(new_sim == "Spring_Summer") / n_observations
simulated_stats[i] <- sim_p
}
sim <- data_frame(p_springsummer = simulated_stats)
sim %>% ggplot(aes(x = p_springsummer)) +
geom_histogram(binwidth = 0.001, colour = "black", fill = "grey") +
xlab("simulated proportion that the incidence happened in spring and summer")
```

```{r echo=FALSE, warning=FALSE, eval = FALSE}
# estimated p_value
sim %>%
filter(p_springsummer >= 0.5035488 | p_springsummer <= 0.4964512) %>%
summarise(p_value = n() / repetitions)
```


## Question 2:
Does the number of autothefts show a significant correlation depending on whether it is during the day or the night?

```{r, echo=FALSE, warning=FALSE}
neighbourhood %>% 
  filter(!is.na(time_of_day)) %>%
  ggplot(aes(x = time_of_day)) + geom_bar(color = "black", fill = "lightblue")+
  ggtitle("Number of Autothefts based on Daytime vs Nighttime")+
  labs(x= "Daytime vs Nighttime", y= "Number of Autothefts")
```

## Question 2 Results

**Null Hypothesis:** The proportion of auto thefts in the daytime and the nighttime are equal (50% happen at night vs 50% happen during the day)

**Method Used:** Simulation with a single variable was run 1000 times

**Results:** Very strong evidence against the Null hypothesis


```{r echo=FALSE, warning=FALSE, eval = FALSE}
#simulation for time of day
n <- 18178
repetitions <- 1000
sim <- rep(NA, repetitions)
set.seed(1071)
for (i in 1:repetitions) {
  sim_time_of_day <- sample(c("Day", "Night"),
       size = n,
       prob = c(0.5, 0.5),
       replace = TRUE) 
  sim_p <- sum(sim_time_of_day == "Day") / n
  sim[i] <- sim_p
}
#build a plot with simulated data
sim <- data_frame(p_day = sim)
sim %>% ggplot(aes(x = p_day)) +
  geom_histogram(binwidth = 0.001, colour = "black", fill = "lightblue") + ggtitle("Simulated Distribution of daytime") + labs(x = "Proportion day time in 18178 auto thefts" )
```

```{r echo=FALSE, warning=FALSE, eval = FALSE}
#results for time of day
neighbourhood %>% filter(!is.na(time_of_day))%>%
  summarise(n(),
            n_daytime = sum(time_of_day == "Day"),
            prop = n_daytime / n())
```

```{r echo=FALSE, warning=FALSE, eval = FALSE}
#p_hat for time of day
6509/(18178)
#     = 0.3580702
#test statistic
      0.5 - (0.3580702 - 0.5)
#     = 0.6419298
#p-value
      sim %>% filter(p_day >= 0.6419298 | p_day <= 0.3580702) %>%
        summarise(p_value = n() / repetitions)
#     = 0
```


## Question 3:
Does the number of autothefts show a significant correlation depending on whether it is during the Spring/Summer or Winter/Fall at midnight?

```{r, echo=FALSE, warning=FALSE}
neighbourhood %>% filter(!is.na(occurrencehour), !is.na(occurrencemonth)) %>% 
  ggplot (aes(x = occurrencehour, y= "N", fill = season))+
  geom_bar(stat="identity")+
  labs(title = "Number of Auto Thefts Depending on the Hour\nand Season", x="Hour", y="Number of Auto Thefts", colour = "Seasons")+
  theme_bw()+
    theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16),
        axis.text.y = element_text(size = 14), axis.title.y = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold"))+
  scale_fill_discrete(name="Seasons", breaks = c("Fall_Winter", "Spring_Summer"), labels=c("Fall and Winter", "Spring and Summer"))
```

## Classification Tree Depending on Midnight Proportions and Seasons
```{r echo=FALSE, warning=FALSE, eval=TRUE}
neighbourhood1 <- neighbourhood %>% filter(!is.na(season)) %>% mutate(hot_or_not = ifelse(season == "Spring_Summer", yes= "Yes", no = "No")) 
tree1 <- rpart(hot_or_not ~ occurrencehour, data = neighbourhood1)
plot(as.party(tree1), gp=gpar(cex=0.8), type="simple")
```

```{r echo=FALSE, warning=FALSE}
p_neighbourhood <- neighbourhood %>% filter(!is.na(season))
  
n_springsummer <- p_neighbourhood %>% filter(season == "Spring_Summer") %>% summarize(n())
n_fallwinter <- p_neighbourhood %>% filter(season == "Fall_Winter") %>% summarize(n())
 
p_springsummer <- p_neighbourhood %>% filter(season == "Spring_Summer") %>% summarize(prop_SS = sum(occurrencehour == 0))
p_fallwinter <- p_neighbourhood %>% filter(season == "Fall_Winter") %>% summarize(prop_FW = sum(occurrencehour == 0))
  
test_stat <- as.numeric(p_springsummer / n_springsummer - p_fallwinter / n_fallwinter)
n_observations <- 18178
repetitions <- 1000
simulated_stats <- rep(NA, repetitions)
set.seed(1071)
for (i in 1:repetitions)
{
 sim <- p_neighbourhood %>%
   mutate(season = sample(season))
p_springsummer <- sim %>% 
  filter(season == "Spring_Summer") %>% 
  summarize(prop_SS = sum(occurrencehour == 0))
p_fallwinter <- sim %>% 
  filter(season == "Fall_Winter") %>% 
  summarize(prop_FW = sum(occurrencehour == 0))
p_diff <- p_springsummer / n_springsummer - p_fallwinter / n_fallwinter
simulated_stats[i] <- as.numeric(p_diff)
}

n <- nrow(neighbourhood1)
set.seed(1071)
training_indices <- sample(1:n, size=round(0.8*n))
train <- neighbourhood1[training_indices,]
test <- neighbourhood1[-training_indices,]
train_tree <- rpart(hot_or_not ~ occurrencehour, data = train)
prediction_matrix <- predict(object = train_tree, newdata = test, type = "class")            
```






## Question 3 Results

- [INSERT RESULTS]


## Results

Present the main results here, in order of importance, related to the questions asked. You might use tables or graphs, or other ways to summarize your results.

## Conclusion

Give your main conclusions here. Follow the order of questions you presented. 

Here you can also mention any additional considerations, concerns, or issues you might have. For example, if the results were unexpected, you can discuss this and perhaps offer possible explanations.

## Acknowledgements (optional)

If you received any help from someone other than your team members you can acknowledge them. For example:   
*The authors would like to thank "TA name" for helpful suggestions and comments that improved the presentation of this poster.*
