---
title: "The Inclination for Auto Thefts in Relation to Time"
author: 'Stephanie Cristea, Nicole Sin, Johnson Vo, Shuyi Zuo, TUT107, Group Number:
  1'
subtitle: Based on Auto Thefts data from Toronto provided by the the Toronto Police
  Service and from the years 2014-2018
output:
  ioslides_presentation: default
  slidy_presentation: default
  transition: faster
  widescreen: yes
---


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# echo=FALSE will stop the code chunk from appearing in the knit document
# warning=FALSE and message=FALSE will stop R messages from appearing in the knit document
library(tidyverse)
library(rpart)
#install.packages("partykit")
library(partykit)
library(knitr)
auto_thefts <- read.csv("auto_thefts.csv")
neighbourhood <- auto_thefts %>% select(Index_, occurrencemonth, occurrencedayofweek, occurrencehour) %>% filter(occurrencemonth != '' & occurrencedayofweek != '') %>% mutate(season = ifelse((occurrencemonth == 'March'| occurrencemonth == 'April'| occurrencemonth == 'May'| occurrencemonth == 'June'| occurrencemonth == 'July'| occurrencemonth == 'August'), yes = "Spring_Summer", no = "Fall_Winter"))%>%
  mutate(week_end_or_day = ifelse(occurrencedayofweek == "Sunday    "| occurrencedayofweek == "Saturday  ", yes = "Weekend", no = "Weekday")) %>%
  mutate(time_of_day= ifelse(occurrencehour >= 18|occurrencehour <= 6, yes = "Night", no= "Day"))
```
## Goals

- Analyze whether time acts as a factor in the number of auto thefts; is there an inclination to when these crimes occur? 
  - Take an intricate look at the effect of the season and the hour on number of auto thefts
  
- Determine peak time when these crimes occur; when the public should take better precautions protecting their cars 

- Determine when police officers should be more prepared for auto-thefts based on when they take place more often


## Questions of Analysis

1. Does the proportion of auto thefts in the spring/summer differ from the proportion during fall/winter?


2. Does the number of auto thefts show a significant correlation depending on whether it is during the day or the night?


3. Does the number of auto thefts show a significant correlation depending on whether it is during the spring/summer or winter/fall at midnight?


# Data Visualization

---

```{r, echo=FALSE, warning=FALSE, fig.align = "center"}
neighbourhood %>% filter(!is.na(occurrencemonth)) %>%  mutate(month = factor(occurrencemonth, 
                    levels = c("January", "February", "March", "April", 
                               "May", "June", "July", "August", 
                               "September", "October", "November", "December"), ordered = TRUE)) %>%
  ggplot(aes(x = month, fill = season)) + geom_bar() + ggtitle("Number of Auto Thefts Depending on the Months \nof the Year") +
  labs (x = "Month", y = "Number of Auto Thefts")+coord_flip()+
  scale_fill_discrete(name="Seasons", breaks = c("Fall_Winter", "Spring_Summer"), labels=c("Fall and Winter", "Spring and Summer"))+
  theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16),
        axis.text.y = element_text(size = 14), axis.title.y = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
```

---


```{r  echo=FALSE, warning=FALSE, fig.align = "center"}
neighbourhood %>% ggplot(aes(x = occurrencehour, fill = time_of_day)) + geom_bar() + ggtitle("Number of Auto Thefts Depending on the Time \nof the Day")+
  labs (x = "Time of the day (by hour)", y = "Number of Auto Thefts")+
  scale_fill_discrete(name="Time of Day")+
  theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16),
        axis.text.y = element_text(size = 14), axis.title.y = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
```


## Data Summary
The following variables were created to address our questions of analysis:

- A seasonal variable (categorizing an auto theft incidence as "Spring/Summer" or "Fall/Winter")
- A time of day variable (categorizing an auto theft incidence as "Day" or "Night")


## Question 1: 
Does the proportion of auto thefts in the spring/summer differ from the proportion during fall/winter?

```{r echo=FALSE, warning=FALSE}
neighbourhood %>%
  filter(!is.na(season))%>%
  ggplot(aes(x = season))+geom_bar(color = "black", fill = "lightblue")+
  ggtitle("Number of Auto Thefts based on Spring/Summer \nand Fall/Winter")+
  labs (x = "Seasons", y = "Number of Auto Thefts")+
  theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16),
        axis.text.y = element_text(size = 14), axis.title.y = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5))+
  scale_x_discrete(labels=c("Spring/Summer", "Fall/Winter"))+
  geom_text(stat="count", aes(label =..count..), vjust = 2, colour = "white")
```



## Question 1 Results

**Null Hypthothesis:** The proportion of auto thefts int he Spring/Summer is the same as the proportion in the Winter/Fall

**Method Used:** Permutation Test

**Results:** Our analysis shows that there is no evidence against our null hypthosis


```{r echo=FALSE, warning=FALSE}
# null hypothesis: the proportion of the incidence happened in spring and summer is 0.5.
# alternative hypothesis: the proportion of the incidence happened in spring and summer is not 0.5.
```


```{r echo=FALSE, warning=FALSE, eval = FALSE}
# test statistic
neighbourhood %>%
  filter(!is.na(season)) %>%
  summarise(n(),
            n_springsummer = sum (season == "Spring_Summer"),
            prop_springsummer = n_springsummer/n())
```

```{r echo=FALSE, warning=FALSE, eval = FALSE}
# simulation of incidence occurance in seasons. 
n_observations <- 18175
repetitions <- 1000
simulated_stats <- rep(NA, repetitions)
set.seed(1071)
for (i in 1:repetitions)
{
new_sim <- sample(c("Spring_Summer", "Fall_Winter"),
size = n_observations, prob=c(0.5,0.5), replace = TRUE)
sim_p <- sum(new_sim == "Spring_Summer") / n_observations
simulated_stats[i] <- sim_p
}
sim <- data_frame(p_springsummer = simulated_stats)
sim %>% ggplot(aes(x = p_springsummer)) +
geom_histogram(binwidth = 0.001, colour = "black", fill = "grey") +
xlab("simulated proportion that the incidence happened in spring and summer")
```

```{r echo=FALSE, warning=FALSE, eval = FALSE}
# estimated p_value
sim %>%
filter(p_springsummer >= 0.5035488 | p_springsummer <= 0.4964512) %>%
summarise(p_value = n() / repetitions)
```


## Question 2:
Does the number of auto thefts show a significant correlation depending on whether it is during the day or the night?

```{r, echo=FALSE, warning=FALSE}
neighbourhood %>% 
  filter(!is.na(time_of_day)) %>%
  ggplot(aes(x = time_of_day)) + geom_bar(color = "black", fill = "lightblue")+
  ggtitle("Number of Auto Thefts based on Daytime \nvs Nighttime")+
  labs(x= "Daytime vs Nighttime", y= "Number of Auto Thefts")+
  theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16),
        axis.text.y = element_text(size = 14), axis.title.y = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5))+
  geom_text(stat="count", aes(label =..count..), vjust = 2, colour = "white")
```

## Question 2 Results

**Null Hypothesis:** The proportion of auto thefts in the daytime and the nighttime are equal (50% happen at night vs 50% happen during the day)

**Method Used:** Permutation Test

**Results:** Our analysis shows that there is very strong evidence against our null hypothesis. 


```{r echo=FALSE, warning=FALSE, eval = FALSE}
#simulation for time of day
n <- 18175
repetitions <- 1000
sim <- rep(NA, repetitions)
set.seed(1071)
for (i in 1:repetitions) {
  sim_time_of_day <- sample(c("Day", "Night"),
       size = n,
       prob = c(0.5, 0.5),
       replace = TRUE) 
  sim_p <- sum(sim_time_of_day == "Day") / n
  sim[i] <- sim_p
}
#build a plot with simulated data
sim <- data_frame(p_day = sim)
sim %>% ggplot(aes(x = p_day)) +
  geom_histogram(binwidth = 0.001, colour = "black", fill = "lightblue") + ggtitle("Simulated Distribution of daytime") + labs(x = "Proportion daytime in 18175 auto thefts" )
```

```{r echo=FALSE, warning=FALSE, eval = FALSE}
#results for time of day
neighbourhood %>% filter(!is.na(time_of_day))%>%
  summarise(n(),
            n_daytime = sum(time_of_day == "Day"),
            prop = n_daytime / n())
```

```{r echo=FALSE, warning=FALSE, eval = FALSE}
#p_hat for time of day
6509/(18175)
#     = 0.3580702
#test statistic
      0.5 - (0.3580702 - 0.5)
#     = 0.6419298
#p-value
      sim %>% filter(p_day >= 0.6419298 | p_day <= 0.3580702) %>%
        summarise(p_value = n() / repetitions)
#     = 0
```


## Question 3:
Does the number of auto thefts show a significant correlation depending on whether it is during the Spring/Summer or Winter/Fall at midnight?

```{r, echo=FALSE, warning=FALSE}
neighbourhood %>% filter(!is.na(occurrencehour), !is.na(occurrencemonth)) %>% 
  ggplot (aes(x = occurrencehour, y= "N", fill = season))+
  geom_bar(stat="identity")+
  labs(title = "Number of Auto Thefts Depending on the Hour\nand Season", x="Hour", y="Number of Auto Thefts", colour = "Seasons")+
  theme_bw()+
    theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16),
        axis.text.y = element_text(size = 14), axis.title.y = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5))+
 scale_fill_discrete(name="Seasons", breaks = c("Fall_Winter", "Spring_Summer"), labels=c("Fall and Winter", "Spring and Summer"))
```

## Classification Tree Depending on Midnight Proportions and Seasons
*Note: "Yes" refers to Spring/Summer months
```{r echo=FALSE, warning=FALSE, eval=TRUE}
neighbourhood1 <- neighbourhood %>% filter(!is.na(season)) %>% mutate(hot_or_not = ifelse(season == "Spring_Summer", yes= "Yes", no = "No")) 
tree1 <- rpart(hot_or_not ~ occurrencehour, data = neighbourhood1)
plot(as.party(tree1), gp=gpar(cex=0.8), type="simple")
```

```{r echo=FALSE, warning=FALSE}
p_neighbourhood <- neighbourhood %>% filter(!is.na(season))
  
n_springsummer <- p_neighbourhood %>% filter(season == "Spring_Summer") %>% summarize(n())
n_fallwinter <- p_neighbourhood %>% filter(season == "Fall_Winter") %>% summarize(n())
 
p_springsummer <- p_neighbourhood %>% filter(season == "Spring_Summer") %>% summarize(prop_SS = sum(occurrencehour == 0))
p_fallwinter <- p_neighbourhood %>% filter(season == "Fall_Winter") %>% summarize(prop_FW = sum(occurrencehour == 0))
  
test_stat <- as.numeric(p_springsummer / n_springsummer - p_fallwinter / n_fallwinter)
n_observations <- 18175
repetitions <- 1000
simulated_stats <- rep(NA, repetitions)
set.seed(1071)
for (i in 1:repetitions)
{
 sim <- p_neighbourhood %>%
   mutate(season = sample(season))
p_springsummer <- sim %>% 
  filter(season == "Spring_Summer") %>% 
  summarize(prop_SS = sum(occurrencehour == 0))
p_fallwinter <- sim %>% 
  filter(season == "Fall_Winter") %>% 
  summarize(prop_FW = sum(occurrencehour == 0))
p_diff <- p_springsummer / n_springsummer - p_fallwinter / n_fallwinter
simulated_stats[i] <- as.numeric(p_diff)
}
n <- nrow(neighbourhood1)
set.seed(1071)
training_indices <- sample(1:n, size=round(0.8*n))
train <- neighbourhood1[training_indices,]
test <- neighbourhood1[-training_indices,]
train_tree <- rpart(hot_or_not ~ occurrencehour, data = train)
prediction_matrix <- predict(object = train_tree, newdata = test, type = "class")            
```


## Question 3 Results

**Null Hypothesis:** The proportion of auto thefts occuring at midnight between spring/summer seasons and fall/winter seasons are the same.

**Methods Used:** Classification Tree & Permutation Test

**Results:** The classification tree predicts the likelihood of crimes occuring at what time and what season; specifically late night crimes are predicted to be spring/summer crimes and daytime crimes are predicted to be fall/winter.


## Results

**From Our Analyses We Can Conclude:**

- The season has little to no influence on the number of auto thefts
- Time of day has an influence on the number of auto thefts
  - More thefts happen at night than during the day

## Conclusion

Police should strengthen prevention of auto thefts especially during the night regardless of season, which has little to no effect on the number of auto thefts. 

**Additional Considerations:**

- Comparing the seasons individually instead of grouping them (Spring Vs Summer vs Fall vs Winter).  

## Conclusion

**Future Considerations:**

- Day of the week occurrence
- Population density
- Neighbourhood age
- Neighbourhood density
- Premise type

## Acknowledgements
The authors would like to thank Vivian Ngo for helpful suggestions and comments that improved the presentation of this poster.
